---
title: "Modern Statistical Computing"
subtitle: "2. ggplot"
author: "David Rossell"
institute: "Pompeu Fabra University"
execute:
  echo: true
format:
  revealjs:
    theme: [default, custom.scss]
    scrollable: true
toc: true
toc-depth: 1
number-sections: false
mouse-wheel: true
code-overflow: scroll
code-line-numbers: false
code-copy: true
cache: true
title-slide-style: pandoc
bibliography: references.bib
---

## Reproducing these lecture notes

Required R packages

```{r, eval=FALSE}
install.packages("dplyr")
install.packages("gapminder")
install.packages("gganimate")
install.packages("ggplot2")
install.packages("ggstatsplot")
install.packages("fivethirtyeight")
install.packages("tidyverse")
install.packages("devtools")
devtools::install_github("rstudio-education/dsbox")
```

Once they're installed, load them.

```{r, warning=FALSE}
library(tidyverse)  #also loads ggplot2
library(dplyr)
library(dsbox)
library(fivethirtyeight)
library(gapminder)
library(gganimate)
library(ggstatsplot)
```


# Exercise

The blog FiveThirtyEight featured this post on hate crimes around the 2016 USA elections

Dataset `hate_crimes` (package `fivethirtyeight`) has this info for each state:

- `avg_hatecrimes_per_100k_fbi`: pre-election number of hate crimes per 100,000 population by the FBI

- `hate_crimes_per_100k_splc`: post-election crimes in 10-day period from SPLC (southern poverty law center)

- education (% adults $\geq$ 25 with high school degree)

- % of people who voted Donald Trump

- Income inequality (Gini index)

Plus variables related to racial diversity, geographic heterogeneity and economic indicators (see `help(hate_crimes)`)

A simple scatterplot shows that FBI-reported hate crimes are lower in states with more Trump voters. 
However, according to the blog Trump voters were not statistically significant, after accounting for inequality and education

To study this issue, use `ggplot` to produce:

1. Scatter-plots of FBI-related crimes vs. Trump voters, vs. Gini index and vs. high-school. Add a regression line to the plot (without standard errors) and label the x- and y-axes

2. Scatter-plot of FBI-related crimes vs. Trump voters, with different panels (facets) for states in each of 4 categories for the Gini index (see next slide)

3. Same as 2, now with panels for 4 categories in the % of high-school education

To help you get started, load the data and create a smaller version with only the relevant variables

```{r, warning=FALSE}
library(fivethirtyeight)
sel= c('state_abbrev','avg_hatecrimes_per_100k_fbi','share_vote_trump','gini_index','share_pop_hs')
hc= hate_crimes[,sel]
names(hc)= c('state','hatecrimes_fbi','votes_trump','gini','hs')
```

For fun, we obtain their pairwise correlations

```{r, warning=FALSE}
round(cor(hc[,-1], use='complete.obs'),3)
```

Finally, we discretize the Gini index and education into 4 groups

```{r}
hc$ginic= cut(hc$gini, breaks=c(0,0.43,0.45,0.47,1))
levels(hc$ginic)= c('Gini <= 0.43','0.43 < Gini <= 0.45','0.45 < Gini <= 0.47','Gini > 0.47')

hc$hsc= cut(hc$hs, breaks=c(0,0.83,0.86,0.9,1))
levels(hc$hsc)= c('hs <= 0.83', '0.83 < hs <= 0.86', '0.86 < hs <= 0.9', 'hs > 0.9')
```


## Exercise

Load the Gini dataset at the course's GitHub site.

```{r}
library(gganimate)
#library(gifski)  #you may need this in Windows and Mac
gini= as_tibble(read.csv("~/github/statcomp/datasets/gini.csv"))
gini
```

1. Do a (non-animated) line plot of Gini vs time, with a different line colour per country

2. Add `transition_reveal(Year)` to add an animation over time





